% mainfile:main.tex
% revised by dnolivieri

\chapter{Community}\label{chap:community}

When working in a free software project managed by the community, it is very important to be in contact with the 
members of the community.  In the next sections, a description of how to communicate with these people and how 
to contribute to specific project is provided.   As a specific case, the focus of this discussion shall focus 
upon \emph{gedit}\cite{website:gedit}.

\section{Contributing}\label{sec:Contributing}

gedit development, as many other projects from \GNOME, relies upon voluntary contributions and everyone is invited to help.

If anybody is interested in contributing to the development of gedit, the principal way for contacting the developers 
is by IRC (see \ref{sec:IRC}) or sending a message to the gedit mailing list (see \ref{sec:Mailing}).

\section{Code Repository}\label{sec:Repository}

gedit is hosted on the \emph{\GNOME Git repository}\footnote{\url{http://git.gnome.org/}} in the 
gedit module\footnote{\url{http://git.gnome.org/browse/gedit}}. To be able to push the changes to this repository, a 
special account is needed that should be requested to the \GNOME accounts team.  To receive it, a reasonable number of patches 
or bugzilla reports should have been submitted by the person interested. See \ref{sec:Bugs} and \ref{sec:Patches}.

\section{Bug reporting}\label{sec:Bugs}

Bugs should be reported to the \emph{\GNOME bug tracking system}\footnote{\url{http://bugzilla.gnome.org/}}, product gedit\footnote{\url{http://bugzilla.gnome.org/browse.cgi?product=gedit}}.

An email address is needed to register before the system can be used to file a new bug report.

There are guidelines\footnote{\url{http://bugzilla.gnome.org/page.cgi?id=bug-writing.html}} for reporting new bugs, if they are followed, it will help to make things more effective. The use of keywords\footnote{\url{http://bugzilla.gnome.org/describekeywords.cgi}} also helps to prioritize on the bugs.

\section{Patches}\label{sec:Patches}

Patches should be submitted to the bug tracking system (see \ref{sec:Bugs}).

If the proposed patch fixes a known bug, it must be added as an attachment to the corresponding bug report. Otherwise, a new bug report should be created that describes the problem that the patch is designed to fix, and then attach the patch to that bug report.

Patches should be generated using the \textit{git format-patch}\footnote{\url{http://www.kernel.org/pub/software/scm/git/docs/git-format-patch.html}} command and should include a descriptive commit message. See the \textbf{HACKING}\footnote{\url{http://git.gnome.org/cgit/gedit/tree/HACKING}} file for more detailed guidelines.

\section{Mailing lists}\label{sec:Mailing}

Subscriptions to the \emph{gedit mailing list}, or change to an existing subscription can be realized by visiting the gedit-list info page\footnote{\url{http://mail.gnome.org/mailman/listinfo/gedit-list}}.

To post a message to all the list members, send an email to \href{mailto:gedit-list@gnome org}{gedit-list (at) gnome org}.

To see the collection of prior postings to the list, the \href{http://mail.gnome.org/archives/gedit-list/}{gedit-list archives}\footnote{\url{http://mail.gnome.org/archives/gedit-list/}} should be visited.

\section{Internet Relay Chat}\label{sec:IRC}

The developers can be found on the \#gedit channel at \url{irc.gnome.org}.

\section{Wiki}\label{sec:Wiki}

A Wiki is provided for drafting, designing and proposing gedit development, available at \url{http://live.gnome.org/Gedit}.
