% mainfile:main.tex
%  revised by dnolivieri
\chapter{Problems found}

During the development of this project, several important technical and strategic problems were encountered. A summary of 
the most important problmes are given here. 

\section*{Memory leaks}

A principal problem resulting from using C as the core programming language, is that memory leaks must be explicitly handled. 
A memory leak can lead to intrinsic problems and produce unexpected crashes and problems. For such isues, the 
\emph{valgrind} utility helps to deal with this problem.

\section*{Adapt to new libraries}

This was the price to pay for being such on the edge. It meant to update the libraries almost every day and adapt gedit to the API changes and some times have to fix the libraries itself.

**** note:  ``being on the edge'' should be described... it is too much jargon...



\section*{Not enough testing}

Due to lack of time, releases had to be made without sufficient testing. This led to angry users, who reported problems thereby resulting 
in immediate subsequent releases within just a few days after the previous release.

\section*{Miss understanding}

The fact that several people are working on the same project from different parts of the world, and all of them working with a non 
native language, produced communication problems at times that led to reimplement or rethink a problem.

\section*{Limitations}

When using a toolkit or a specific library, great care and consideration must be taken when creating a new design 
based upon the ideal behavior of this library or toolkit.    This can lead to problems, since in fact that library 
may be far from ideal.
